A Fast and Tight Heuristic for A* in Road Networks

### 2021-01-27 paper/ch_potentials.tex
grammar fix, comments, wordsmithing ALT footnote

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
We study exact, efficient and practical algorithms for route planning in large road networks.
Routing applications often requires integrating the current traffic situation, planning ahead with traffic predictions for the future, respecting forbidden turns, and many other features depending on the exact application.
While Dijkstra's algorithm can be used to solve these problems, it is too slow for many applications.
A* is a classical approach to accelerate Dijkstra's algorithm.
A* can support many extended scenarios without much additional implementation complexity.
However, A*'s performance depends on the availability of a good heuristic that estimates distances.
Computing tight distance estimates is a challenge on its own.
On road networks, shortest paths can also be quickly computed using hierarchical speedup techniques.
They achieve speed and exactness but sacrifice A*'s flexibility.
Extending them to certain practical applications can be hard.
In this paper, we present an algorithm to efficiently extract distance estimates for A* from Contraction Hierarchies (CH), a hierarchical technique.
We call our heuristic CH-Potentials.
Our approach allows decoupling the supported extensions from the hierarchical speed-up technique.
Additionally, we describe A* optimizations to accelerate the processing of low degree nodes, which often occur in road networks.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
The past decade has seen a plethora of research on route planning in large street networks~\cite{bdgmpsww-rptn-16}.
Routing a user through a road network can be formalized as the shortest path problem in weighted graphs.
Nodes represent intersections.
Roads are modeled using edges.
Edges are weighted by their traversal times.
The problem can be solved with Dijkstra's algorithm~\cite{d-ntpcg-59}.
Unfortunately, on continental sized networks, it is too slow for many applications.
Thus, speed-up techniques have been developed.
One popular example are Contraction Hierarchies~(CH)~\cite{gssv-erlrn-12}.
They have been used successfully in many real world applications.
A CH exploits the inherent hierarchy of road networks.
In a preprocessing step, additional shortcut edges are inserted, which allow skipping unimportant parts of the network at query time.
Another popular example is Multi-Level-Dijkstra~(MLD)~\cite{swz-umlgt-02} also known as CRP~\cite{dgpw-crprn-13}.
It is also used in practice~\cite{bingblog}.
MLD also uses shortcut edges.
Both approaches achieve speed-ups of three orders of magnitude over Dijkstra's algorithm.

@@

Truck routing differs from car routing due to night driving bans and other restrictions.
In~\cite{kswz-erptd-p-20}, a preliminary version of CH-Potentials\footnote{
In~\cite{kswz-erptd-p-20}, CH-Potentials are used as a blackbox referring to an early ArXiv preprint~\cite{strasser2019perfect} of ours.
Our submitted paper is the finished version of the ArXiv preprint.
The submitted paper does not describe any of the contribution of \cite{kswz-erptd-p-20}.
} is used for such a scenario.
The work considers time-dependent blocked edges and waiting at parking locations.
Further, a trade-off between arrival time and route quality is considered.

@@

Table~\ref{tab:building_blocks} depicts the performance of A* with different heuristics and optimizations.
We compare CH-Potentials to three other heuristics.
First, the Zero heuristic where $h(x)=0$ for all nodes $x$.
This corresponds to using Dijkstra's algorithm.
Second, we compare against our own implementation\footnote{
We use 16 landmarks generate with the avoid strategy~\cite{gw-cppsp-05} and activate all during every query.
%We store all landmark distances consecutive for each node.
%This increases cache efficiency when reading many distances for the same node.
%We do not disable some landmarks during the query, but use all landmarks all the time.
Our ALT implementation is uni-directional, as a bidirectional search creates problems for some settings, such as predicted traffic.
%Also, we do not employ bidirectional search because we want to evaluate and compare the building block CH-Potentials in its simplest and most flexible form.
} of ALT~\cite{gw-cppsp-05}.
Finally, we compare against a hypothetical \emph{Oracle-A*} heuristic.
This heuristic has instant access to a shortest distance array with respect to $w_\ell$, i.e. it is the most efficient heuristic possible in our model.
We fill this array before each query using a reverse Dijkstra search from the target node.
Thus, the reported running times of Oracle-A* do \emph{not} account for any heuristic evaluation.
Comparing against Oracle-A* allows us to measure the overhead of the heuristic evaluation of CH-Potentials.
Also, no other heuristic, which only has access to the preprocessing weights, can be faster than Oracle-A*.

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
We study exact, efficient and practical algorithms for route planning in large road networks.
Routing applications often require integrating the current traffic situation, planning ahead with traffic predictions for the future, respecting forbidden turns, and many other features depending on the exact application.
While Dijkstra's algorithm can be used to solve these problems, it is too slow for many applications.
A* is a classical approach to accelerate Dijkstra's algorithm.
A* can support many extended scenarios without much additional implementation complexity.
However, A*'s performance depends on the availability of a good heuristic that estimates distances.
Computing tight distance estimates is a challenge on its own.
On road networks, shortest paths can also be quickly computed using hierarchical speedup techniques.
They achieve speed and exactness but sacrifice A*'s flexibility.
Extending them to certain practical applications can be hard.
In this paper, we present an algorithm to efficiently extract distance estimates for A* from Contraction Hierarchies (CH), a hierarchical technique.
We call our heuristic CH-Potentials.
Our approach allows decoupling the supported extensions from the hierarchical speed-up technique.
Additionally, we describe A* optimizations to accelerate the processing of low degree nodes, which often occur in road networks.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
The past decade has seen a plethora of research on route planning in large street networks~\cite{bdgmpsww-rptn-16}.
Routing a user through a road network can be formalized as the shortest path problem in weighted graphs.
Nodes represent intersections.
Roads are modeled using edges.
Edges are weighted by their traversal times.
The problem can be solved with Dijkstra's algorithm~\cite{d-ntpcg-59}.
Unfortunately, on continental sized networks, it is too slow for many applications.
Thus, speed-up techniques have been developed.
One popular example are Contraction Hierarchies~(CH)~\cite{gssv-erlrn-12}.
They have been used successfully in many real world applications.
A CH exploits the inherent hierarchy of road networks.
In a preprocessing step, additional shortcut edges are inserted, which allow skipping unimportant parts of the network at query time.
Another popular example is Multi-Level-Dijkstra~(MLD)~\cite{swz-umlgt-02} also known as CRP~\cite{dgpw-crprn-13}.
It is also used in practice~\cite{bingblog}.
MLD also uses shortcut edges.
% ich haette das at least eigentlich schon gerne drin. CH erreicht wesentlich mehr als 3 groessenordnungen
Both approaches achieve speed-ups of three orders of magnitude over Dijkstra's algorithm.

@@

Truck routing differs from car routing due to night driving bans and other restrictions.
In~\cite{kswz-erptd-p-20}, a preliminary version of CH-Potentials\footnote{
In~\cite{kswz-erptd-p-20}, CH-Potentials are used as a blackbox referring to an early ArXiv preprint~\cite{strasser2019perfect} of ours.
Our submitted paper is the finished version of the ArXiv preprint.
% der punkt macht fuer mich keinen Sinn.
% Dass wir keine der Contributions von kswz-erptd-p-20 beschreiben ist ja offensichtlich, sieht ja jeder in dem Abschnitt.
% Wichtiger finde ich, dass kswz-erptd-p-20 nix von uns beschreibt.
% Das meinte ich zumindest mit dem Satz den du geloescht hast.
% Aber der jetzt folgende Satz sagt nur, dass unser paper hier nix von kswz-erptd-p-20 beschreibt.
The submitted paper does not describe any of the contribution of \cite{kswz-erptd-p-20}.
} is used for such a scenario.
The work considers time-dependent blocked edges and waiting at parking locations.
Further, a trade-off between arrival time and route quality is considered.

@@

Table~\ref{tab:building_blocks} depicts the performance of A* with different heuristics and optimizations.
We compare CH-Potentials to three other heuristics.
First, the Zero heuristic where $h(x)=0$ for all nodes $x$.
This corresponds to using Dijkstra's algorithm.
Second, we compare against our own implementation\footnote{
We use 16 landmarks generated with the avoid strategy~\cite{gw-cppsp-05} and activate all during every query.
Our ALT implementation is uni-directional.
In this work, we do not consider bidirectional search as it creates problems for some settings, such as predicted traffic.
} of ALT~\cite{gw-cppsp-05}.
Finally, we compare against a hypothetical \emph{Oracle-A*} heuristic.
This heuristic has instant access to a shortest distance array with respect to $w_\ell$, i.e. it is the most efficient heuristic possible in our model.
We fill this array before each query using a reverse Dijkstra search from the target node.
Thus, the reported running times of Oracle-A* do \emph{not} account for any heuristic evaluation.
Comparing against Oracle-A* allows us to measure the overhead of the heuristic evaluation of CH-Potentials.
Also, no other heuristic, which only has access to the preprocessing weights, can be faster than Oracle-A*.
